A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion

### 2017-03-24 manuscript.tex
restarted working on manuscript

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Vestibular Contributions to Spatial Perspective Transformations: From
Sensory Inference To Imagined Self-Motion}

@@

  \authornote{
    \newcounter{author}
    Andrew W. Ellis \& Fred W. Mast, Department of Psychology. AE and FM
    planned and conducted the research; AE and FM analyzed the data and
    wrote the manuscript. Code for simulations is available at
    \url{https://osf.io/u26mq}.

@@

  \abstract{Recent research has indicated that the vestibular network is involved in
various cognitive tasks involving mental spatial transformations.
However, the nature of this involvement remains unclear. Here, we
provide a computational framework for vestibular cognition and discuss
this in the context of perspective transformations, which require a
simulated rotation of the self. We explore how the ability to model the
effects of body rotation using vestibular input might enable the ability
to imagine body rotations. We approach this problem by extending
previous models of vestibular sensory processing to include the ability
to model the sensory consequences of self-initiated movements, and
discuss how this can be further extended to allow mentally simulated
rotations of the self. We \ldots{} Predicting the sensory input is key
to understanding imagined rotations. Mental imagery goes beyond
prediction, rather, it involves the use of counterfactual queries in a
probabilistic model.}
  \keywords{probabilistic graphical models, particle filter, embodied cognition,
vestibular cognition, spatial perspective taking, mental imagery \\

@@

  \usepackage{caption}

@@

% pandoc-xnos: macro to create a caption without a prefix
\makeatletter
\long\def\@makenoprefixcaption#1#2{
  \vskip\abovecaptionskip
  \sbox\@tempboxa{#2}
  \ifdim \wd\@tempboxa >\hsize
    #2\par
  \else
    \global \@minipagefalse
    \hb@xt@\hsize{\hfil\box\@tempboxa\hfil}
  \fi
  \vskip\belowcaptionskip}
\makeatother

% pandoc-fignos: save original macros
\makeatletter
\let\@oldmakecaption=\@makecaption
\let\oldthefigure=\thefigure
\let\oldtheHfigure=\theHfigure
\makeatother

% pandoc-fignos: environment disables figure caption prefixes
\makeatletter
\newcounter{figno}
\newenvironment{no-prefix-figure-caption}{
  \let\@makecaption=\@makenoprefixcaption
  \renewcommand\thefigure{x.\thefigno}
  \renewcommand\theHfigure{x.\thefigno}
  \stepcounter{figno}
}{
  \let\thefigure=\oldthefigure
  \let\theHfigure=\oldtheHfigure
  \let\@makecaption=\@oldmakecaption
  \addtocounter{figure}{-1}
}
\makeatother

% pandoc-xnos: cleveref fakery
\newcommand{\plusnamesingular}{}
\newcommand{\starnamesingular}{}
\newcommand{\xrefname}[1]{\protect\renewcommand{\plusnamesingular}{#1}}
\newcommand{\Xrefname}[1]{\protect\renewcommand{\starnamesingular}{#1}}
\providecommand{\cref}{\plusnamesingular~\ref}
\providecommand{\Cref}{\starnamesingular~\ref}
\providecommand{\crefformat}[2]{}
\providecommand{\Crefformat}[2]{}

% pandoc-xnos: cleveref formatting
\crefformat{figure}{Figure~#2#1#3}
\Crefformat{figure}{Figure~#2#1#3}

@@

Adopting the spatial perspective of another person is generally
considered an essential cognitive ability, and plays a vital role in our
ability to determine what another person can see and in order to
coordinate joint actions (Creem-Regehr, Gagnon, Geuss, \& Stefanucci,
2013, Pezzulo, Iodice, Ferraina, \& Kessler (2013)). Many studies have
also pointed out that perspective taking is of fundamental importance
for language comprehension and communication (Beveridge \& Pickering,
2013), and social cognition in general (Deroualle \& Lopez, 2014).

A number of recent studies have demonstrated that the vestibular system,
which deals with sensory signal relating to movements of the head, is
involved in such spatial transformations (Lenggenhager, Lopez, \&
Blanke, 2008, Deroualle, Borel, Devèze, \& Lopez (2015), Falconer \&
Mast (2012), Gardner, Stent, Mohr, \& Golding (2016)). This is
demonstrated in the form of interference effects between mental
transformations and concurrent sensory stimulation. However, to date,
there has little attempt to understand the involvement of the vestibular
system in higher-level cognitive abilities from a computational
perspective. In this paper, we focus specifically on the role of the
vestibular system for spatial perspective transformations, and we draw
the link between probabilistic computations performed by the vestibular
system in the service of real-time interaction with the world, and the
brain's ability to run mental simulations for the purpose of adopting
the perspective of another person.

Consider the use of deictic pronouns, such as \enquote{this} or
\enquote{that} and \enquote{left} or \enquote{right}. These terms only
assume a their meaning when placed within a certain frame of reference
(FOR). For example, in a dialogue between two speakers, the statement
\enquote{I would like that one}, or \enquote{I would like the one on the
right} can only be understood under the assumption that the deictic
terms \enquote{that} and \enquote{right} are used in an egocentric FOR.
In addition, both interlocutors must be able to access the other's FOR
with respect to an external FOR. In other words: In order to give
instructions, the speaker must be able to mentally assume the
addressee's position in space. A further example that requires a similar
spatial transformation is when two people want to perform a coordinated
action, such as lifting a heavy object. This requires having knowledge
of the other's potential actions in order to adjust one's own actions,
and this again requires knowledge of the other's position with an
external FOR. A judgement about what kind of actions the other person
can perform requires an embodied simulation using a representation of
one's own body schema (Pezzulo et al., 2013). Giving an instruction to
another person thus requires running an embodied simulation in order to
infer what kind of motor commands the other person should perform
\textbf{{[}Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod,
2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo,
2011a,b{]}}. This embodied simulation is performed using one's own body
schema. Communicating this information to the other actor requires
knowing the position of the addressee's FOR relative to the external
reference frame, and the difference between the speaker and the
addressee's egocentric reference frames. Perspective taking thus can be
conceived of as a aligning a representation of one's body with that of
another person in order to access spatial information from a viewpoint
other than one's own egocentric viewpoint. In order to achieve this
alignment, one must perform translations and rotations of one's own
egocentric reference frame. Processing translations and rotations of the
head and body is the core domain of the vestibular system, and thus is
of particular interest for research into mental spatial transformations.
Spatial transformations in 6 dimensions (translation along 3 axes, and
rotations around 3 axes) are highly complex, and non-linear, and we
propose that the brain adopts the strategy of simulating motion of the
body through space in order achieve a desired target position. This
effectively amounts to a running a simulation of dynamical system,
consisting of simulated limb movements and the resulting sensory
consequences. \autoref{fig:spt-schematic} (a) shows a simple scenario in
which a speaker wants to give an addressee, standing opposite, verbal
instructions about an action to perform on an external object. The
speaker must determine the most suitable action, given the addressee's
position relative to the external object, and in order to achieve this
goal, the speaker must align himself with the addressee, so that he can
use his own body-centred reference frame to quickly simulate the best
action. In this example, alignment with the addressee's reference frame
consists of a forwards translation and a rotation around the vertical
axis. This can be described as an imagined movement, in which the goal
is achieve a specific position.

\begin{figure}
\centering
\includegraphics[width=0.50000\textwidth]{diagrams/fig-1-common-process.pdf}
\caption{Perspective taking and imagined self-motion involve spatial
transformations of a representation of the self. (a) If a speaker's goal
is to take the perspective of an addressee, he must first solve the
inverse problem of determining the appropriate action, and then run a
forwards simulation in order to perform a translation and rotation of
his own egocentric reference frame. These operations can be viewed as
imagined movements.(b) Focussing on imagined rotation around the
vertical axis allows us to simplify the problem, and highlights the fact
that perspective taking and mental imagery involve common
computations.\label{fig:spt-schematic}}
\end{figure}

In order to do this, the speaker must first determine what type of
movement to perform, and the direction, velocity and duration that lead
to the desired outcome. This means that in order to know what to
simulate, the speaker must first work backwards from the desired
outcome, and infer the actions that are most likely to lead to that
outcome. This is known in the field of motor control as the inverse
problem (Wolpert \& Kawato, 1998). In order to achieve the desired
position, the motor commands can then be simulated, and the expected
sensory consequences can be estimated; this is known as the forward
problem. \xrefname{Figure}\cref{fig:spt-schematic} (b) shows that if the
goal is to simply simulate the perceptual experience of rotation around
the body's vertical (yaw) axis, rather than project a representation of
one's body onto a target, the problem is simplified; the actor can
perform forward inference without the need to determine the optimal
motor action. As demonstrated by Penny, Zeidman, \& Burgess (2013),
these computations can be performed in a common probabilistic model.
Interestingly, this has been discussed in the context of reinforcement
learning, and planning in humans has been described as reverse inference
(Botvinick \& Toussaint, 2012). Further, (Pickering \& Clark, 2014)
discuss the relationship between control-theoretic notions of
inverse/forward models, known collectively as internal models, and and
more general concepts of a generative model in a hierarchical Bayesian
framework.

Performing a simulation of body movements and estimating the sensory
consquences amounts to simulating the behaviour of a dynamical system;
this requires a generative model of the system. The aim of this paper is
to specifically investigate the generative model that may be used for
mental spatial transformations. The ability to run embodied simulations
is often claimed to be grounded in the relevant sensorimotor circuits,
and we claim that the vestibular system is ideally suited to investigate
this connection, because the generative models used by the vestibular
system for dynamic sensory inference have been extensively investigated
using approaches from optimal control theory, and more recently, dynamic
Baysian inference. The vestibular system faces a number of challenges in
interpreting the sensory measurements provided by the semi-circular
canals (SCC) and otolith organs. On the one hand, these signals are very
noisy, and ambiguous (MacNeilage, Ganesan, \& Angelaki, 2008); the
otoliths respond both to translations and tilt of the head relative to
gravity, and in order to disambiguate between translation and tilt, the
brain must perform sophisticated computations at the level of the brain
stem. It has become clear that the nature of these computations is
probabilistic, and that the brain makes use of strong prior beliefs in
order to interpret these sensory signals. A further challenge faced by
the vestibular system is that it needs to distinguish between sensory
signals that are the result of active motion, and those resulting from
external forces (Cullen, Brooks, \& Sadeghi, 2009, Cullen (2014)). The
ability to distinguishing re-afferent from ex-afferent sensory signals
is an essential capability of all organisms that move actively (Crapse
\& Sommer, 2008).

Furthermore, the apparent connection between mental imagery and
predictive processing has repeatedly been alluded to (Moulton \&
Kosslyn, 2009, Clark (2012), Gambi \& Pickering (2015)), and Grush
(2004) claims that mental imagery is performed by emulator circuits
implementing a forward model. These ideas can also be addressed from a
computational point of view in the context of dynamic Bayesian
inference, where there are multiple roles for sensory predictions, over
different time scales. This was recently discussed by (Tian \& Poeppel,
2012) in the context of imagined speech.

We proceed by introducing recent models of Bayesian inference for
real-time, dynamic sensory processing of vestibular data, based on
particle filters (Laurens \& Droulez, 2007, Karmali \& Merfeld (2012)),
and we reframe these as probabilistic graphical models. This allows us
to investigate the generative model separately from the algorithm used
for inference, and we then extend the basic model for sensory inference
in order to provide the ability to model the consequences of
self-initiated, active movements. Our aims are to:

@@

@@

Whereas early models of vestibular sensory processing were inspired by
control theory (Borah, Young, \& Curry, 1988, Merfeld, Zupan, \& Peterka
(1999)), several recent studies have described vestibular sensory
processing as dynamic Bayesian inference (Laurens \& Droulez, 2007,
MacNeilage et al. (2008), Karmali \& Merfeld (2012)). Selva \& Oman
(2012) provide a recent review of the relationship between optimal
observer models and models which employ Kalman filtering, and show that
these are largely equivalent. Both Karmali \& Merfeld (2012) and Laurens
\& Droulez (2007) described vestibular sensory inference in terms of
particle filtering, which is a particular inference algorithm that can
be used to perform inference in state-space models. (Doucet, Godsill, \&
Andrieu, 2000, Doucet \& Johansen (2009)). Using their model, Laurens \&
Droulez (2007) were able to simulate perceptual responses to
centrifugation and off-vertical axis rotations.

A further well-known phenomenon of vestibular perception is velocity
storage, which is defined as the slower decay of perceptual and
oculomotor responses, relative to the SCC afferent responses (Raphan,
Matsuo, \& Cohen, 1979), \textbf{{[}Robinson 1977{]}}. This is generally
seen as evidence for the existence of an internal model in the
processing of vestibular signals. Karmali \& Merfeld (2012) focussed on
one-dimensional angular velocity during yaw rotations; their model was
able to explain the phenomenon of velocity storage as a consequence of
particle filtering, with the time constant of velocity storage depending
on the variability of the sensory afferent signals.

@@

We proceed by providing a brief introduction to the concept of
state-space models, and particle filtering, which is a particular
algorithm that can perform inference in state-space models. While Kalman
filters are restricted to linear Gaussian State-space models, particle
filtering can be applied to arbitrary nonlinear models. State-space
models can be seen as special cases of more general Bayesian
hierarchical models (Bishop, 2006; Murphy, 2012). Following the same
approach as Penny et al. (2013), we first describe the generative model
used for sensory inference as a dynamic probabilistic graphical model.

A state-space is special case of a dynamic latent variable model which
is suitable for modelling dynamical system in terms of latent,
unobservable processes and noisy observations. As an example, consider
the case where we want to model the velocity and position during a
leftward head turn. While the SCC measures angular acceleration, the
afferent signals are thought to be proportional to the angular velocity
of the head with a frequency band of approximately 0.1 -- 10 Hz, which
constitutes the range of natural head movements (Carriot, Jamali,
Chacron, \& Cullen, 2014; Grabherr, Nicoucar, Mast, \& Merfeld, 2008).
The measurements are observations of the true, unobservable angular
velocity; however, they are contaminted by Gaussian noise, which is
determined by characteristics of the sensory organs.
\autoref{fig:head-turn-data} shows a sequence of noisy measurements
during a two-second turn of the head. The acceleration generating the
movement is sinusoidal, given for each time step \(t\) by
\(Asin(2\pi t/T)\), where \(A\) is the amplitude of the movement,
\(T = 1/f\) is the duration of the movement, and \(f\) is the frequency.
This acceleration results in a peak velocity \(\omega_{peak} = AT/\pi\)
and a final angular position \(\theta_{final} = AT^2/2\pi\).

\begin{no-prefix-figure-caption}

\begin{figure}
\centering
\includegraphics{../generated-figures/head-turn-data-1.pdf}
\caption{\label{fig:head-turn-data}Sensory measurements obtained during a
short head turn. The goal is to infer the latent angular velocity and
position, based on the noisy measurements. \label{fig:head-turn}}
\end{figure}

\end{no-prefix-figure-caption}

The key idea is that we can write model the dynamics of a system by the
evolution of the latent state variables in discrete time steps as a
Markov process, where the state at each time point \(t\) depends only on
the previous state at time \(t-1\). This is referred to as the process
model. Estimating the velocity \(\omega\) and position \(\theta\) in
discrete time therefore requires writing the kinematic equations as a
set of first order difference equations.

\[\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t\]
\[\omega_{t} = \omega_{t-1}\]

>>>>

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={A Computational Perspective On Spatial Perspective Transformations: From Sensory Inference To Imagined Self-motion},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{A Computational Perspective On Spatial Perspective Transformations: From
Sensory Inference To Imagined Self-motion}

@@

  \authornote{
    \newcounter{author}
    Andrew W. Ellis \& Fred W. Mast, Department of Psychology. AE and FM
    planned and conducted the research; AE and FM performed the simulations
    and wrote the manuscript. Code for simulations is available at
    \url{https://osf.io/u26mq}.

@@

  \abstract{Mental spatial transformations are considered to be an essential
cognitive ability. Recent research has indicated that the vestibular
network is involved in such mental spatial transformations. However, the
nature of this involvement remains unclear. We propose a computational
framework for the involvement of the vestibular system in cognition and
discuss this in the context of perspective transformations, which
require a simulated rotation of the self. We explore how the ability to
anticipate the sensory consequences of self-initiated actions might
underlie the ability to imagine body rotations. We approach this problem
by extending previous models of vestibular sensory inference to include
the ability to model the sensory consequences of self-initiated
movements, and discuss how this can be further extended to allow
mentally simulated rotations of the self. Our model is able to explain
various source of interference between simulations of self-rotation and
sensory inference.}
  \keywords{probabilistic graphical models, particle filter, embodied cognition,
vestibular cognition, spatial perspective taking, mental imagery, mental
simulation \\

@@

@@

@@

Adopting the spatial perspective of another person is an essential
cognitive ability, and plays a vital role in our ability to determine
possible actions of another person, and in order to coordinate joint
actions (Creem-Regehr, Gagnon, Geuss, \& Stefanucci, 2013, Pezzulo,
Iodice, Ferraina, and Kessler (2013)). Perspective taking is also
considered to be of fundamental importance for language comprehension
and communication (Beveridge \& Pickering, 2013), and social cognition
in general (Deroualle \& Lopez, 2014).\\
 A number of recent studies have demonstrated that the vestibular
system, which processes sensory signal relating to angular and linear
movements of the head, is involved in such spatial transformations
(Lenggenhager, Lopez, \& Blanke, 2008, Deroualle, Borel, Devèze, and
Lopez (2015), Falconer and Mast (2012), Gardner, Stent, Mohr, and
Golding (2016)). This is demonstrated in the form of interference
effects between mental transformations and concurrent sensory
stimulation. To date, however, there has little attempt to understand
the involvement of the vestibular system in higher-level cognitive
abilities in terms of the computational mechanisms. Consider the use of
deictic pronouns, such as \enquote{this} or \enquote{that} and
\enquote{left} or \enquote{right}. These terms only assume a meaning
relative to a certain frame of reference (FOR). For example, in a
dialogue between two speakers, the statement \enquote{I would like that
one}, or \enquote{I would like the one on the right} can only be
understood if both interlocutors are able to access the other's FOR with
respect to an external FOR. A further example that requires a similar
spatial transformation is when two people want to perform a coordinated
action, such as lifting a heavy object. This requires having knowledge
of the other's potential actions in order to adjust one's own actions,
and this again requires knowledge of the other's position within an
external FOR. A judgement about what kind of actions the other person
can perform requires an embodied simulation using a representation of
one's own body schema (Pezzulo et al., 2013) and giving an instruction
to another person thus requires running an embodied simulation in order
to infer what kind of motor commands the other person should perform
\textbf{{[}Blakemore and Decety, 2001; Wolpert et al., 2003; Jeannerod,
2006; Pezzulo et al., 2007, 2013; Dindo et al., 2011; Pezzulo,
2011a,b{]}}. But this again requires information about the addressee's
FOR relative to the external reference frame, and the difference between
the speaker and the addressee's egocentric reference frames. Perspective
taking can be conceived of as a aligning a representation of one's body
with that of another person in order to access spatial information from
a viewpoint other than one's own egocentric viewpoint. In order to
achieve this alignment, one must perform translations and rotations of
one's own egocentric reference frame. Processing translations and
rotations of the head and body is the core domain of the vestibular
system, and thus is of particular interest for research into mental
spatial transformations. Spatial transformations in 6 dimensions
(translation along 3 axes, and rotations around 3 axes) are highly
complex, and non-linear, and it has been proposed that the brain adopts
the strategy of simulating motion of the body through space in order
achieve a desired target position {[}CITATION{]}. This effectively
amounts to a running a simulation of dynamical system, consisting of
simulated limb movements and the resulting sensory consequences. Figure
\ref{fig:spt-schematic} (a) shows a simple scenario in which a speaker
wants to tell an addressee, standing opposite, to pick something up from
the floor. The speaker must determine the most suitable action by
embodied simulation, given the addressee's position relative to the
external object, and in order to achieve this goal, the speaker must
align himself with the addressee, so that he can use his own
body-centred FOR. In this example, alignment with the addressee's
reference frame consists of a forwards translation and a rotation around
the vertical axis. This can be described as an imagined movement, in
which the goal is achieve a specific position.

\begin{figure}
\centering
\includegraphics{diagrams/fig-1-common-process.pdf}
\caption{\label{fig:spt-schematic}Perspective taking and imagined
self-motion involve spatial transformations of a representation of the
self. (a) If a speaker's goal is to take the perspective of an
addressee, he must first solve the inverse problem of determining the
appropriate action, and then run a forwards simulation in order to
perform a translation and rotation of his own egocentric reference
frame. These operations can be viewed as imagined movements.(b)
Focussing on imagined rotation around the vertical axis allows us to
simplify the problem, and highlights the fact that perspective taking
and mental imagery involve common computations.}
\end{figure}

In order to do this, the speaker must first determine what type of
movement to perform. This means that in order to know what to simulate,
the speaker must first work backwards from the desired outcome, and
infer the actions that are most likely to lead to that outcome (Hamrick
\& Griffiths, 2014). In the field of motor control, this is known as the
inverse problem (Wolpert \& Kawato, 1998). In order to achieve the
desired position, the motor commands can then be simulated, and the
expected sensory consequences can be estimated; this is known as the
forward problem. Figure \ref{fig:spt-schematic} (b) shows that if the
goal is to deliberatively simulate the perceptual experience of rotation
around the body's vertical (yaw) axis, rather than project a
representation of one's body onto a target, the problem is simplified;
the actor can perform forward inference without the need to determine
the optimal motor action. As demonstrated by Penny, Zeidman, and Burgess
(2013), these computations can be performed in a common probabilistic
model. Interestingly, this has been discussed in the context of
reinforcement learning, and planning in humans has been described as
reverse inference (Botvinick \& Toussaint, 2012). Further, (Pickering \&
Clark, 2014) discuss the relationship between control-theoretic notions
of inverse/forward models, known collectively as internal models, and
and more general concepts of a generative model in a hierarchical
Bayesian framework. Simulating the behaviour of a dynamical system using
a forward model requires a generative model of the system.

The aim of this paper is to specifically investigate the generative
model that may be used for mental spatial transformations. The ability
to run embodied simulations is often claimed to be grounded in the
relevant sensorimotor circuits, and we claim that the vestibular system
is ideally suited to investigate this connection, because the generative
models used by the vestibular system for dynamic sensory inference have
been extensively investigated using approaches from optimal control
theory, and more recently, dynamic Bayesian inference (Oman, CITATIONS,
MacNeilage, Ganesan, \& Angelaki, 2008).

The vestibular system faces a number of challenges in interpreting the
sensory measurements provided by the semi-circular canals (SCC) and
otolith organs. On the one hand, these signals are very noisy, and
ambiguous; the otoliths respond both to translations and tilt of the
head relative to gravity, and in order to disambiguate between
translation and tilt, the brain must perform sophisticated computations.
The emerging picture is that the nature of these computations is
probabilistic, and that the brain makes use of strong prior beliefs in
order to interpret these sensory signals {[}Mittelstaedt, etc{]}. A
further challenge faced by the vestibular system is that it needs to
distinguish between sensory signals that are the result of active
motion, and those resulting from external forces (Cullen, Brooks, \&
Sadeghi, 2009, Cullen (2014)). The ability to distinguishing re-afferent
from ex-afferent sensory signals is an essential capability of all
organisms that move actively (Crapse \& Sommer, 2008).

Furthermore, the apparent connection between mental imagery and
predictive processing has repeatedly been alluded to (Moulton \&
Kosslyn, 2009, Clark (2012), Gambi and Pickering (2015)), and Grush
(2004) claims that mental imagery is performed by emulator circuits
implementing a forward model. These ideas can also be addressed from a
computational point of view in the context of dynamic Bayesian
inference, where there are multiple roles for sensory predictions, over
different time scales. This was recently discussed by (Tian \& Poeppel,
2012) in the context of imagined speech. The putative link between
predictive processing and higher cognitive processing has a firm basis
in predictive processing accounts of sensory inference {[}Friston{]}.
The predictive processing framework has been applied to almost all
sensory modalities {[}CITATIONS: pain, proprioception, auditory ,
visual, etc.{]}, and the picture that has emerged in recent years is
that the brain employs sophisticated inference algorithms in order to
\enquote{interpret} the stream of incoming sensory data. The brain seems
to be actively engaged in formulating hypotheses, and testing these on
sensory data; this has been extended to the concept of active inference
{[}Friston{]}, in which an agent attempts to either change it's internal
model of world, or to intervene on the world in order to fulfill sensory
predictions.

We claim that the vestibular modality is particularly well-suited to
investigating predictive processing at multiple hierarchical scales,
from the lowest levels of sensory inference to higher levels, where
goals, intentions and beliefs of the organism become relevant. A case in
point is the study by Wertheim, Mesland, and Bles (2001), who provided
evidence that participants' perceptions of tilt induced by vestibular
stimulation are shaped to a large extent by the beliefs about possible
motions they formed prior to taking part in the experiment.

Interference effects between higher-level cognitive processes, such as
SPT and mental imagery, and sensory inference, are thought to occur as a
result of shared circuitry (Deroualle \& Lopez, 2014, Nigmatullina et
al. (2015)). However, the precise nature of such interactions, and what
exactly might be shared, is left unspecified. At first glance, it appear
unlikely that such interference should be possible, as low levels of
vestibular processing appear to be cognitively impenetrable
{[}CITATION{]}. In particular, Nigmatullina et al. (2015) found that the
onset of the vestibulo-ocular reflex (VOR) was earlier when the
direction of an imagined self-rotation was congruent with the direction
of a subsequently experienced rotation. The VOR, however, must be
susceptible to the influence of intentions and goals; for example, if a
person intends to fixate a target during a movement, the VOR should be
cancelled. Thus, seemingly reflexive, and impenetrable, motor actions
require a large amount of flexibility based on top-down information.
Even accepting the fact that low level sensorimotor processing may be
more cognitively penetrable than previously thought, the fact that
imagining a rotation of the self should have an effect on a person's
ability to detect rotations, or to discriminate between directions of
rotation, still requires an explanation.

\begin{quote}
NOTE: we need to include all studies demonstrating interference,
especially in the direction cognition -\textgreater{} inference
(Nigmatullina effect)
\end{quote}

Ellis and Mast (2017) recently claimed that the types of interference
effects might arise as a result of cognitive processing exerting an
influence on circuits performing sensory inference via shared
higher-level priors. In particular, the interference described by
Nigmatullina et al. (2015) looks like an effect of a bias in the sensory
discrimination task. Although vestibular perceptual thresholds have been
extensively investigate for yaw rotations (Grabherr, Nicoucar, Mast, \&
Merfeld, 2008; Merfeld, 2011), the methods used for estimating
thresholds require unbiased responding by the participants. There have
no investigations of biasing effects on yaw rotations.

In this paper, we pursue the claim that interference should be
investigated as a bias, and we attempt to explain the grounding of
higher level cognitive abilities in the computations performed by
low-level sensory circuits in the service of dynamic real-time
interactions with the world.

Kwisthout, Bekkering, and Rooij (2016) make the point that predictive
processing accounts are situated at low levels in the processing
hierarchy, usually at the level of sensory inference or at the level
describing kinematics.

In this paper, we address this issue by starting from a model developed
for low-level sensory inference, incorporating basic abilities to
predict sensory consequences of movement based on knowledge about the
kinematics, and extending this to account for interactions between
thought processes, operationalized as mental simulations of sensorimotor
experience, and sensory processing.

Battaglia, Hamrick, and Tenenbaum (2013) addressed the idea that the
higher level cognitive ability to perform mental rotation are grounded
in the circuits performing sensory inference. We claim that if
vestibular cognition is grounded in sensory inference, and that this is
to be experimentally studied by analyzing interactions between thought
and sensory inference, then it is vital that this be carried within the
constraints of a theoretical (computational) framework. A suitable
formalism for this is the framework of probabilistic graphical models
(Pearl, 1988, Koller and Friedman (2009)), and within this framework,
interactions between thought and perception can be modelled as effects
of higher level priors and hyperpriors on inference at a computational
level. Investigating inference at the level of algorithmic
implementations might provide further insight.

Probabilistic graphical models are able to represent hierarchically
structured knowledge (Kwisthout et al., 2016, Griffiths, Vul, and
Sanborn (2012)), which is required for Bayesian models of cognition
(Tenenbaum, Kemp, Griffiths, \& Goodman, 2011, Lucas and Griffiths
(2010)). This approach has been used to model concept learning (Kemp,
Perfors, \& Tenenbaum, 2007), inductive biases in reasoning, in
cognitive models of development (A. Perfors, Tenenbaum, Griffiths, \&
Xu, 2011), language (Goodman \& Frank, 2016) and social cognition
(Baker, Saxe, \& Tenenbaum, 2009).

In this paper, we focus specifically on the role of the vestibular
system for spatial perspective transformations, and we draw the link
between probabilistic computations performed by the vestibular system,
and the brain's ability to run mental simulations for the purpose of
adopting the perspective of another person.

We proceed by introducing recent models of Bayesian inference for
real-time, dynamic sensory processing of vestibular data, based on
particle filters (Laurens \& Droulez, 2007, Karmali and Merfeld (2012)),
and we re-frame these as probabilistic graphical models. This allows us
to investigate the generative model separately from the algorithm used
for inference, and we then extend the basic model for sensory inference
in order to focus specifically on the ability to model the consequences
of self-initiated, active movements. This is an implementation of the
idea that information about motor commands are projected to sensory
areas by means of an efference copy, based on which a forward model then
computes the sensory consequences. Our aims are to:

@@

Following Karmali and Merfeld (2012), we focus on one specific type of
head movement, a rotation of the head about the yaw axis, and discuss
this in the context of the probabilistic computations that are performed
when the goal is to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  infer motion caused by external events
\item
  infer the consequences of one's own actions.
\end{enumerate}

@@

Whereas early models of vestibular sensory processing were inspired by
control theory (Borah, Young, \& Curry, 1988, Merfeld, Zupan, and
Peterka (1999)), several recent studies have described vestibular
sensory processing as dynamic Bayesian inference (Laurens \& Droulez,
2007, MacNeilage et al. (2008), Karmali and Merfeld (2012)). Selva and
Oman (2012) provide a recent review of the relationship between optimal
observer models and models that employ Kalman filtering, and show that
these are largely equivalent. Both Karmali and Merfeld (2012) and
Laurens and Droulez (2007) described vestibular sensory inference in
terms of particle filtering, which is a particular algorithm that can be
used to perform inference in state-space models. (Doucet, Godsill, \&
Andrieu, 2000, Doucet and Johansen (2009)). Using their model, Laurens
and Droulez (2007) were able to simulate perceptual responses to
centrifugation and off-vertical axis rotations. A well-known phenomenon
of vestibular perception is velocity storage, which is defined as the
slower decay of perceptual and oculomotor responses, relative to the SCC
afferent responses (Raphan, Matsuo, \& Cohen, 1979), \textbf{{[}Robinson
1977{]}}. This is generally seen as evidence for the existence of an
internal model in the processing of vestibular signals. Karmali and
Merfeld (2012) focused on one-dimensional angular velocity during yaw
rotations; their model was able to explain the phenomenon of velocity
storage as a consequence of particle filtering, with the time constant
of velocity storage depending on the noise of the sensory afferent
signals.

@@

We proceed by providing a brief introduction to the concept of
state-space models, and particle filtering, which is a particular
algorithm that can be used to perform inference in state-space models.
State-space models can be seen as special cases of more general dynamic
probabilistic graphical models (Bishop, 2006; Murphy, 2012), and are
suitable for modelling the time-varying behaviour of a system in terms
of a latent, unobservable process and noisy measurements.

\subsubsection{Dynamics/kinematics}\label{dynamicskinematics}

For example, consider trying to estimate the velocity and position
during a leftward head turn. In this case, the latent process provides a
model of the kinematics of the system, and the measurement model
describes how the afferent SCC signals are generated, dependent on the
kinematics.

The SCC signals are considered to be proportional to the angular
velocity of the head within a frequency band of approximately 0.1 -- 10
Hz, which constitutes the range of natural head movements (Carriot,
Jamali, Chacron, \& Cullen, 2014; Grabherr et al., 2008). These
measurements, however, are contaminated by Gaussian noise, which is
determined by characteristics of the sensory organs. Figure
\ref{fig:head-turn-data} shows a sequence of noisy measurements during a
two-second leftward turn of the head. The acceleration generating the
movement is sinusoidal, given for each time step \(t\) by
\(Asin(2\pi t/T)\), where \(A\) is the amplitude of the movement,
\(T = 1/f\) is the duration of the movement, and \(f\) is the frequency.
This acceleration results in a half-cycle cosine velocity profile, with
a peak velocity \(\omega_{peak} = AT/\pi\) and a final angular position
\(\theta_{final} = AT^2/2\pi\). Particle filtering provides a means of
filtering out the noise inherent in the measurements, and allows
estimation of evolution of the unobservable, latent state variables.

\begin{figure}
\centering
\includegraphics{generated-figures/head-turn-data-1.pdf}
\caption{\label{fig:head-turn-data}Sensory measurements obtained during a
short leftward head turn. The goal is to infer the latent angular
velocity and position, based on the noisy measurements.}
\end{figure}

\subsubsection{State-space notation}\label{state-space-notation}

The dynamics of the system can be modelled by the evolution of the 3-d
latent state variable\footnote{A state-space model reduces an
  n\textsuperscript{th} order difference equations (in discrete time) to
  n first-order difference equations.}, representing the position and
its derivatives, velocity and acceleration, as a first-order Markov
process, where the state at each time point \(t\) depends only on the
previous state at time \(t-1\). This is referred to as the process
model. Estimating the position \(\theta\), velocity \(\omega\) and
acceleration \(\alpha\) in discrete time requires writing the kinematic
equations as a set of first order difference equations.
